Reduced invariants from cuspidal maps

### 2018-01-15 draft_merged.tex
final review

\documentclass[11pt]{amsart}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
%\usepackage{showlabels}
\usepackage{amsthm}

@@

\newcommand{\M}[4]{\overline{\pazocal M}_{#1,#2}(#3,#4)}
\newcommand{\Mone}[3]{\overline{\pazocal M}^{(1)}_{#1}(#2,#3)}
\newcommand{\PP}{\mathbb P}
\renewcommand{\k}{\mathbf k}
\newcommand{\OO}{\mathcal O}
\renewcommand{\to}{\rightarrow}
\newcommand{\pr}{\rm{pr}}
\newcommand{\Aaff}{\mathbb A}
\newcommand{\A}{\pazocal A}
\newcommand{\B}{\pazocal B}
\newcommand{\X}{\pazocal X}
\newcommand{\XP}{\pazocal{X\!P}}
\newcommand{\tXP}{\widetilde{\pazocal{X\!P}}}
\renewcommand{\L}{\mathcal L}
\newcommand{\hL}{\widehat{\mathcal L}}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\hC}{\widehat{\mathcal C}}
\newcommand{\Z}{\pazocal Z}
\newcommand{\Zp}{\pazocal Z^p}
\newcommand{\tZ}{\widetilde{\pazocal Z}}
\newcommand{\tZp}{\widetilde{\pazocal Z}^p}
\newcommand{\MM}{\mathfrak M}
\newcommand{\pP}{\mathfrak P}
\newcommand{\hP}{\widehat{\mathfrak P}}
\newcommand{\hM}{\widehat{\mathfrak M}}
\newcommand{\hrM}{\widehat{\rm{M}}}
\newcommand{\oM}{\overline{\pazocal M}}
\newcommand{\tM}{\widetilde{\pazocal M}}
\newcommand{\R}{\operatorname{R}}
\newcommand{\Gm}{\mathbb{G}_{\rm{m}}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\PtoM}{\lambda}
\newcommand{\vir}[1]{[#1]^{\rm{vir}}}
\newcommand{\virloc}[1]{[#1]^{\rm{vir}}_{\rm{loc}}}
\newcommand{\dvr}{\Delta}

@@

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{defprop}[thm]{Definition-Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem*{teo*}{Theorem}
\newtheorem{ipotesi}{ipotesi}
\newtheorem*{nota}{Nota}
\newtheorem{claim}{Claim}
\newtheorem{question}[thm]{Question}

@@

\title{Reduced GW invariants from cuspidal curves}
\author{L.Battistella, F.Carocci, C.Manolache}
\date{\today}
\begin{document}

@@

We prove that the $1$-stabilisation morphism, replacing elliptic tails
with cusps, is well-defined at the level of genus $1$ weighted-stable
curves. We use this to endow a substack of $\oM_1(\PP^4,d)$ with a
virtual class that ignores the boundary contribution of maps with a
contracted elliptic tail. We thus prove that Vakil-Zinger's reduced
invariants of the quintic threefold can be computed as Smyth-Viscardi's
$1$-stable (i.e. cuspidal) invariants. To do so we adapt Jun Li's
techniques of $p$-fields and local equations for the moduli space.

@@

\maketitle

\section*{Introduction}

@@

Recall that in genus zero $\M{0}{n}{\PP^r}{d}$ is a smooth stack of the expected dimension $r-3+d(r+1)+n$. Furthermore the rational Gromov-Witten theory of a complete intersection $i\colon X\hookrightarrow \PP^r$ of degree $(l_1,\ldots,l_k)$ can be determined as a twisted theory for $\PP^r$ and thus computed via localisation: for any split vector bundle $E=\oplus_{i} \mathcal O_{\PP^r}(l_i)$ the sheaf $\pi_*f^*E$ is a vector bundle on $\M{0}{n}{\PP^r}{d}$, where we have denoted by $\pi$ the projection from the universal curve and by $f$ the universal stable map.
%\bcd
%\mathcal C_{0,n}(\PP^r,d)\ar[r,"f"]\ar[d,"\pi"] & \PP^r%\times \M{0}{n}{\PP^r}{d} \\
%\M{0}{n}{\PP^r}{d} &
%\ecd
It is a matter of functoriality of virtual fundamental classes \cite{KON} \cite{CKL} \cite{KKP} that the following \emph{hyperplane property} holds
%$$\bigsqcup_{\beta\in A_1(X)\colon i_*(\beta)=d}\M{0}{n}{X}{\beta}\hookrightarrow\M{0}{n}{\PP^r}{d}$$
%then the following equality holds:, which is expected from the interpretation of Chern classes as zero loci of sections (if $X$ is cut by a generic section $s\in H^0(\PP^r,E)$, then the induced section $\tilde s=\pi_*f^*(s)$ is also generic):
$$j_*[\bigsqcup_{\beta\in A_1(X)\colon i_*(\beta)=d}\M{0}{n}{X}{\beta}]^{\operatorname{vir}}=c_{\operatorname{top}}(E)\cap[\M{0}{n}{\PP^r}{d}].$$
%This, so-called ``hyperplane property'', allows us to compute GW invariants of complete intersections as twisted invariants of projective space, where computations can be done via localisation.

For $g\geq 1$ the moduli space $\M{g}{n}{\PP^r}{d}$ is neither irreducible, nor pure dimensional, and the sheaf $\pi_*f^*\mathcal O_{\PP^r}(l)$ is not a vector bundle. The enumerative meaning of the invariants is spoilt by degenerate contributions from curves of lower genera, represented by boundary components of the moduli space where a subcurve of positive genus is contracted.

The situation in genus $1$ is better understood: the geometry of $\M{1}{n}{\PP^r}{d}$ has been widely studied since \cite{Vre}. Contrary to higher genus, the closure of the locus of maps the source of which is a smooth elliptic curve is an irreducible component of $\M{1}{n}{\PP^r}{d}$, which is usually called the \emph{main component}.
Its boundary points were characterised in \cite{Vre} as follows: $[f\colon C\to \PP^r]\in\oM_1(\PP^r,d)$ is smoothable, i.e. it can be made into the special fiber of a $1$-dimensional family the generic fiber of which is a non-constant map from a smooth elliptic curve, if and only if either $f$ is not constant on the minimal genus $1$ subcurve $E$ of $C$, or the image of the tangent vectors to the rational tails are linearly dependent in $T_{f(E)}\PP^r.$  In Section~\ref{sec:section1} we give a new proof of this fact~\ref{prop:smoothability} using Hu-Li's local equations for the moduli space; furthermore we provide a different characterisation of the smoothable maps as those that factor through a non-degenerate map from a Cohen-Macaulay genus~$1$ singularity \ref{prop:smoothability2}, which we classify extending an argument due to Smyth.
%One of the reasons for this (in fact the only reason in genus one) is that there are maps from reducible curves with a genus $g$ core that is contracted and some rational tails that carry all the degree; if $f\colon C\to \PP^r$ is such a map, then $H^1(C,f^*\mathcal O_{\PP^r}(1))\neq 0$, as can be easily seen from the normalisation sequence.

From the enumerative point of view, it would be ideal to isolate the contribution to the invariants given by the main component.
In an attempt to extract this information hidden within Kontsevich's moduli space, 
Vakil and Zinger managed to desingularise the main component by performing a sequence of blow-ups on an underlying space of weighted curves \cite{VZpreview} \cite{VZ} (see also \cite{HL}). The end result $\widetilde{\pazocal M}_{1,n}(\PP^r,d)^{\rm{main}}$ is a smooth and irreducible stack on which $\tilde\pi_*\tilde f^*\mathcal O_{\PP^r}(l)$ is a vector bundle. This fact allowed them to define a notion of \emph{reduced invariants} for complete intersections in $\PP^r$ by mimicking the hyperplane property of the genus $0$ theory. %The idea is that blowing up successively the locus of degree zero elliptic curve with one rational tail (\emph{elliptic tails}), then with two (\emph{elliptic bridges}), etc. and taking the fiber product
%\bcd
%\widetilde{\mathcal M}_{1,n}(\PP^r,d)\ar[r]\ar[d] & \M{1}{n}{\PP^r}{d}\ar[d] \\
%\widetilde {\mathfrak M}_{1,n}^{\operatorname{wt}=d} \ar[r] & \mathfrak M_{1,n}^{\operatorname{wt}=d}
%\ecd
%the strict transform of the main component becomes smooth and meets the rest of the components in normal crossing fashion (the \'{e}tale local model being the union of two linear subspaces of different dimensions in affine space). Furthermore, the sheaves $\tilde\pi_*\tilde f^*\mathcal O_{\PP^r}(l)$ contain a vector subbundle of rank $dl$ supported on the strict transform of the main component. This paved the way to the definition of reduced invariants.

\begin{dfn}
Let $X$ be a complete intersection of degree $(l_1,\ldots,l_k)$ in $\PP^r$. The genus $1$ reduced invariants of $X$ are defined by integrating against: $$ c_{\rm{top}}(\tilde\pi_*\tilde f^*(\oplus_{i=1}^k\mathcal O_{\PP^r}(l_i)))\cap[\widetilde{\pazocal M}_{1,n}(\PP^r,d)^{\rm{main}}].$$
\end{dfn}

Recently there has been a different attempt to discard some of the boundary components. Based on Smyth's work on the birational geometry of $\oM_{1,n}$ \cite{SMY1,SMY2}, Viscardi \cite{VISC} has introduced a series of \emph{alternate compactifications} of the moduli space of maps from smooth elliptic curves. Compared to Kontsevich's space of stable maps, in Viscardi's space $\oM_{1,n}^{(m)}(X,\beta)$ maps that contract a genus $1$ curve with at most $m$ rational tails are destabilised, and replaced by maps from $l$-elliptic singularities with $l\leq m$.

@@

\begin{thm}
Let $X_5\subseteq\PP^4$ be a generic smooth quintic threefold. Then:
\[
N_1^{\rm{red}}(X_5,d)=N_1^{\rm{cusp}}(X_5,d),
\]
where:
\[N_1^{\rm{red}}(X_5,d):=\deg \left(c_{\rm{top}}(\tilde\pi_*\tilde f^*(\mathcal O_{\PP^4}(5)))\cap [\widetilde{\pazocal M}_{1}(\PP^4,d)^{\rm{main}}]\right)\]
and
\[N_1^{\rm{cusp}}(X_5,d):=\deg [\Mone{1}{X_5}{d}]^{\rm{vir}}.
\]
\end{thm}

Below we will suppress the subscript $5$ from $X_5$. The insight for this theorem comes from the following formula, the proof of which was first given by J. Li and A. Zinger \cite{LZ} in the symplectic category:
\begin{equation}\label{eq:VakilZinger}
N_1(X,d)=N_1^{\rm{red}}(X,d)+\frac{1}{12}N_0(X,d).
\end{equation}
The first proof of Li-Zinger's formula in the algebro-geometric setting is due to H.-W. Chang and J. Li in \cite{CL}. Their splitting of the intrinsic normal cone makes it patent that the genus $0$ contribution to the right hand side only comes from the boundary component with a single rational tail of degree $d$. This suggests that removing such a component would provide a more direct approach to reduced invariants for the quintic threefold. This is precisely what Viscardi's space of $1$-stable maps does.

%To prove the result we essentially need to understand how the various component of the moduli space contribute to the invariants, so we have to look at the algebraic techniques that  have been used to prove~\eqref{eq:VakilZinger}.

%Indeed, after the symplectic proof, Li-Zinger's formula was proved . The latter is only the final a series of papers where they developed the theory of stable maps with $p$-fields and applied it together with the techniques of cosection-localised virtual classes \cite{KLcosection}. $p$-fields provide a way around the fact that we cannot control the geometry of $\oM_1(X,d)$ and there is no compatible perfect obstruction theory relative to the inclusion $\oM_1(X,d)\hookrightarrow \oM_1(\PP^4,d)$; indeed Chang and Li managed to include the unwanted part of the obstruction theory in the moduli space, and by these means endow a well-understood space (very similar to $\oM_1(\PP^4,d)$) with an obstruction theory counting genus $1$ curves in the quintic.
The proof of our main theorem follows from adapting the techniques of Chang and Li to Viscardi's moduli spaces. Here is an outline of the proof.

%Alternatively, if one does not want to use the theory of stable maps with fields, the idea would be to follow the alternative proof of~\eqref{eq:VakilZinger} recently proposed by \cite{preprint}. In that case, they still consider the Hu-Li blow up of the moduli space of weighted curve and they then look at $\tM_1(X)\hookrightarrow \tM_1(\PP^4) $, where $X$ is the quintic threefold and study directly the cone of $\tM_1(X).$ Following~\cite{preprint} steps, we can attack directly $\tM_1(X)^{(1)}$ and show that only the main component has a non trivial contribution to the invariants.

%In both cases the comparison with the genus~1 invariants comes from the following comparison result.
%We show the following:
We first consider a moduli space of $1$-stable maps with $p$-fields $\Mone{1}{\PP^4}{d}^p$. In the work of Chang and Li, $p$-fields provide a way around the fact that we cannot control the geometry of $\oM_1(X,d)$ and there is no compatible perfect obstruction theory relative to the inclusion $\oM_1(X,d)\hookrightarrow \oM_1(\PP^4,d)$. By introducing the moduli space of maps with $p$-fields the authors manage to include the unwanted part of the obstruction theory into the moduli space, and to endow a well-understood space (indeed very similar to $\oM_1(\PP^4,d)$) with an obstruction theory counting genus $1$ curves in the quintic. Paralleling their proof, i.e. using cosection localised virtual cycles \cite{KLcosection} and the deformation to the normal cone $(X_5\subseteq \PP^4)\rightsquigarrow(X_5\subseteq N_{X/\PP^4})$, we see that:
\begin{prop}
$\deg\virloc{\Mone{1}{\PP^4}{d}^p}=\deg\vir{\Mone{1}{X}{d}}.$
\end{prop}
Indeed it is enough to notice that all the constructions in \cite{CLpfields} carry over to the setting of a family of Gorenstein (not necessarily nodal) curves.

We then need the following comparison morphism.
\begin{prop}
There exists a well-defined $1$-stabilisation morphism at the level of weighted-stable curves:
\[
\MM_{1,n}^{\rm{wt,st}}\to \MM_{1,n}^{\rm{wt,st}}(1)
\]
\end{prop}
The fiber product
\[
\Z:=\MM_{1,n}^{\rm{wt,st}}\times_{\MM_{1,n}^{\rm{wt,st}}}\Mone{1}{\PP^4}{d}
\]
is a substack of $\oM_1(\PP^4,d)$ avoiding the boundary component with one rational tail of degree $d$.

It is harder to compare the fiber product
\[
\Zp:=\MM_{1,n}^{\rm{wt,st}}\times_{\MM_{1,n}^{\rm{wt,st}}}\Mone{1}{\PP^4}{d}
\]
with $\oM_1(\PP^4,d)^p$. Yet $\Zp$ is endowed with a cosection-localised virtual class counting $1$-stable maps to the quintic threefold (by cosection-localised virtual pullback). The same splitting arguments as in \cite{CLpfields} allow us to evince a main contribution which coincides with Vakil-Zinger's reduced invariants. On the other hand a dimensional computation shows that this is the only enumeratively meaningful component.

\subsection*{Future directions}

\subsection*{Outline of the paper}

\subsection*{Notations and conventions} We work over an algebraically closed field $\k$ of characteristic $0$.

Throughout the paper we shall have fixed a homogeneous polynomial $\mathbf w\in\k[x_0,\ldots,x_4]_5$ of degree $5$ such that $X=V(\mathbf w)\subseteq\PP^4$ is a generic smooth quintic threefold. Furthermore let us consider a fixed positive integer $d$ determining all the homology classes $\beta=d[\ell]\in A_1(\PP^4)$, line bundle weights, etc. that we shall take into account.

@@

Let $\MM_{1,n}$ denote the algebraic stack of prestable $n$-marked curves of arithmetic genus $1$. In the following definitions, by \emph{special point} we mean the preimage of a node or a marking in the normalisation.
\begin{dfn}
Let $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$ be the stack of \emph{weighted-stable} curves of genus $1$ with $n$ markings and total weight $d$: every geometric point of $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$ represents a connected, reduced, nodal, projective curve of arithmetic genus $1$ with $n$ distinct smooth markings and an integer-valued function $\rm{wt}$ on the dual graph. Such a function assumes only nonnegative values, the sum of which on all the vertices of the dual graph is $d$; $\rm{wt}$ is compatible with the specialisation maps, and we further impose the following stability condition: every $p_a=0$ component of weight $0$ has at least three special points (every $p_a=1$ component of weight $0$ has at least one special point).
\end{dfn}
There is an \'{e}tale non-separated morphism $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}\to\mathfrak M_{1,n}$ and the stability condition is such that the forgetful map $\M{1}{n}{\PP^r}{d}\to\mathfrak M_{1,n}$ factors through $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$, the weight coming from the degree of the map to $\PP^r$.
We shall denote by $\pi\colon \cC\to \MM_{1,n}$ the universal curve (and use similar notation for its pullbacks). We shall also consider the universal Picard stack $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}$ of $\pi\colon \cC\to \MM_{1,n}$ with universal line bundle $\mathcal{L}$, subject to the stability condition that $\omega_\pi^{\rm{log}}\otimes\mathcal L^{\otimes 3}$ is $\pi$-ample, and the requirement that the total degree of $\mathcal L$ on the fibers of $\cC$ is $d$. Let $\PtoM$ denote the projection $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}\to \mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$, where the weight comes from the degree of the line bundle. The forgetful map $\rm M=\M{1}{n}{\PP^r}{d}\to\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$ factors through $\PtoM$ by declaring $\mathcal L_{\rm M}:=f^*\OO_{\PP^r}(1)$. We shall often use shorthand notation $\MM$ for $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$ and $\mathfrak P$ for $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}$.

\begin{dfn}
Let $\mathfrak M_{1,n}^{\operatorname{wt}=d,\text{st}}(1)$ be the stack of \emph{at worst cuspidal} projective connected, reduced, $n$-marked, projective curves of arithmetic genus $1$ that are weighted-stable with total weight $d$, i.e. the weight is nonnegative, and every $p_a=0$ component of weight $0$ has at least three special points (every $p_a=1$ component of weight $0$ has at least \emph{two} special points).
\end{dfn}
\begin{remark}
 It follows from the miniversal deformation of the cusp that being at worst cuspidal (i.e. having only nodes and cusps as singolarity) is an open condition on the base of any family of curves.
\end{remark}

We shall denote by $\hat{\pi}\colon \widehat{\cC}\to \MM_{1,n}(1)$ the universal curve (and use similar notation for its pullbacks). We shall also consider the universal Picard stack $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}(1)$ of $\widehat{\cC}\to \mathfrak M_{1,n}(1)$ with universal line bundle $\widehat{\mathcal{L}}$ subject to similar stability conditions to the one spelled out for nodal curves. We shall often use shorthand notation $\widehat{\MM}$ for $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}(1)$ and $\widehat{\mathfrak P}$ for $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}(1)$. Let $\hat{\PtoM}$ denote the projection $\widehat{\mathfrak P}\to \widehat{\MM}$.

>>>>

\documentclass[11pt]{amsart}
\usepackage[english]{babel}
\usepackage[margin=1.6in]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
%\usepackage{showlabels}
\usepackage{amsthm}

@@

\newcommand{\M}[4]{\overline{\pazocal M}_{#1,#2}(#3,#4)}
\newcommand{\Mone}[3]{\overline{\pazocal M}^{(1)}_{#1}(#2,#3)}
\newcommand{\PP}{\mathbb P}
\renewcommand{\k}{\mathbf k}
\newcommand{\OO}{\mathcal O}
\renewcommand{\to}{\rightarrow}
\newcommand{\pr}{\rm{pr}}
\newcommand{\Aaff}{\mathbb A}
\newcommand{\A}{\pazocal A}
\newcommand{\B}{\pazocal B}
\newcommand{\X}{\pazocal X}
\newcommand{\XP}{\pazocal{X\!P}}
\newcommand{\tXP}{\widetilde{\pazocal{X\!P}}}
\renewcommand{\L}{\mathcal L}
\newcommand{\hL}{\widehat{\mathcal L}}
\newcommand{\C}{\mathfrak C}
\newcommand{\cC}{\mathcal C}
\newcommand{\hC}{\widehat{\mathcal C}}
\newcommand{\Z}{\pazocal Z}
\newcommand{\Zp}{\pazocal Z^p}
\newcommand{\tZ}{\widetilde{\pazocal Z}}
\newcommand{\tZp}{\widetilde{\pazocal Z}^p}
\newcommand{\oZp}{\overline{\Z}^{p,\rm{gst}}}
\newcommand{\MM}{\mathfrak M}
\newcommand{\pP}{\mathfrak P}
\newcommand{\hP}{\widehat{\mathfrak P}}
\newcommand{\hM}{\widehat{\mathfrak M}}
\newcommand{\hrM}{\widehat{\rm{M}}}
\newcommand{\oM}{\overline{\pazocal M}}
\newcommand{\tM}{\widetilde{\pazocal M}}
\newcommand{\R}{\operatorname{R}}
\newcommand{\Gm}{\mathbb{G}_{\rm{m}}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\PtoM}{\lambda}
\newcommand{\vir}[1]{[#1]^{\rm{vir}}}
\newcommand{\virloc}[1]{[#1]^{\rm{vir}}_{\rm{loc}}}
\newcommand{\dvr}{\Delta}

@@

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{defprop}[thm]{Definition-Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem*{teo*}{Theorem}
\newtheorem*{matteo*}{Main Theorem}
\newtheorem{ipotesi}{ipotesi}
\newtheorem*{nota}{Nota}
\newtheorem{claim}{Claim}
\newtheorem{question}[thm]{Question}

@@

\title{Reduced invariants from cuspidal maps}
\author{L.Battistella, F.Carocci, C.Manolache}
\date{\today}

\setcounter{tocdepth}{1}
\begin{document}

@@

We provide a modular interpretation of Vakil-Zinger reduced invariants of the quintic threefold in genus $1$. Inspired by Li-Zinger's formula, we prove they coincide with $1$-stable invariants, arising from Smyth-Viscardi's alternate compactification where elliptic tails are replaced by cusps.

@@

\maketitle
\tableofcontents

\section{Introduction}

@@

Contrary to the genus $0$ case, $\M{1}{n}{\PP^r}{d}$ has many irreducible components of different dimension \cite{Vre}, and the enumerative geometry of complete intersections $X\subseteq \PP^r$ is hardly accessible through this space, since the hyperplane property does not hold \cite{KKP}.

On the other hand the locus of maps from a smooth elliptic curve is irreducible, and its closure is an irreducible component of $\M{1}{n}{\PP^r}{d}$ which is called the \emph{main component}.

R. Vakil and A. Zinger \cite{VZ} produced a desingularisation $\widetilde{\pazocal M}_{1,n}(\PP^r,d)^{\rm{main}}$ of the main component, on which $\tilde\pi_*\tilde f^*\mathcal O_{\PP^r}(l)$ is a vector bundle. This is achieved through an iterated blow-up construction, which unfortunately makes the modular meaning of this space unclear; though an interpretation has recently been given in terms of log structures \cite{RSPW}. Vakil and Zinger introduced then a notion of \emph{reduced invariants} for complete intersections in $\PP^r$: if $X$ is a complete intersection of degree $(l_1,\ldots,l_k)$, its genus $1$ reduced invariants are defined by integrating against:
\[ c_{\rm{top}}(\tilde\pi_*\tilde f^*(\oplus_{i=1}^k\mathcal O_{\PP^r}(l_i)))\cap[\widetilde{\pazocal M}_{1,n}(\PP^r,d)^{\rm{main}}].\]
Recently there has been a different attempt to discard some of the boundary components. Based on D.I. Smyth's work on the birational geometry of $\oM_{1,n}$ \cite{SMY1,SMY2}, M. Viscardi \cite{VISC} has introduced a series of \emph{alternate compactifications} of the moduli space of maps from smooth elliptic curves, the first instance of which is $\oM_{1,n}^{(1)}(X,\beta)$, where elliptic tails are made unstable and cuspidal singularities are allowed in the source curve.

@@

\begin{matteo*}\label{thm:main}
Let $X_5\subseteq\PP^4$ be a generic smooth quintic threefold. Then:
\[
N_1^{\rm{red}}(X_5,d)=N_1^{\rm{cusp}}(X_5,d),
\]
where:
\[N_1^{\rm{red}}(X_5,d):=\deg \left(c_{\rm{top}}(\tilde\pi_*\tilde f^*(\mathcal O_{\PP^4}(5)))\cap [\widetilde{\pazocal M}_{1}(\PP^4,d)^{\rm{main}}]\right)\]
and
\[N_1^{\rm{cusp}}(X_5,d):=\deg [\Mone{1}{X_5}{d}]^{\rm{vir}}.
\]
\end{matteo*}
The insight for this result comes from the following formula, first proved by J. Li and A. Zinger \cite{LZ} in the symplectic category:
\begin{equation*}
N_1(X_5,d)=N_1^{\rm{red}}(X_5,d)+\frac{1}{12}N_0(X_5,d).
\end{equation*}
An algebro-geometric proof of Li-Zinger's formula is due to H.-W. Chang and J. Li in \cite{CL}. Their method reveals that the genus $0$ contribution to the right hand side only comes from the boundary component with a single rational tail. This suggests that removing such a component would provide a more direct approach to reduced invariants for the quintic threefold. This is precisely what Viscardi's space of $1$-stable maps does.

Here is a rough sketch of how we prove our result. We first show:
\begin{teo*}
There exists a well-defined $1$-stabilisation morphism at the level of weighted-stable curves:
\[
\MM_{1,n}^{\rm{wt,st}}\to \MM_{1,n}^{\rm{wt,st}}(1)
\]replacing elliptic tails of weight $0$ with cusps.
\end{teo*}
We may then use this morphism to construct a substack of $\oM_1(X_5,d)$ that has no component with contracted elliptic tails:
\[\MM_{1}^{\rm{wt,st}}\times_{\MM_{1}^{\rm{wt,st}}(1)}\Mone{1}{X_5}{d},\]
and endow it with a virtual class with the same enumerative content of $\vir{\Mone{1}{X_5}{d}}$. However it is not easy to compare the latter directly with Vakil-Zinger's reduced invariants.

Therefore we follow in Chang and Li's footsteps: we introduce the moduli space of $1$-stable maps with $p$-fields, which has the advantage of having a simpler geometry and admits a cosection-localised virtual class with the same degree as $\vir{\Mone{1}{X_5}{d}}$, see \cite{CLpfields}. We construct:
\[\MM_{1}^{\rm{wt,st}}\times_{\MM_{1}^{\rm{wt,st}}(1)}\Mone{1}{\PP^4}{d}^p,\]
and perform a desingularisation of it via the study of local equations \cite{HL}, followed by an analysis of a splitting of the intrinsic normal cone \cite{CL}, which delivers the theorem.

On our way there, we review Vakil's characterisation of maps lying in the boundary of $\oM_1(\PP^r,d)^{\rm{main}}$ and prove it holds using Hu-Li's equations. We also review and extend Smyth's work on genus $1$ singularities, and rephrase smoothability as follows.

\begin{teo*}
A map $[f]\in\oM_1(\PP^r,d)$ contracting the minimal genus $1$ subcurve is smoothable if and only if it factors through a map from a genus $1$ singularity that does not contract any component.
\end{teo*}

%\subsection*{Future directions}

\subsection{Outline of the paper}
In \S~\ref{sec:section1} we review some classical results about the irreducible components of $\oM_1(\PP^r,d)$ and local equations for this moduli stack in a smooth ambient space. We recall Smyth's $m$-elliptic singularities and Viscardi's alternate compactifications  of the space of maps. We discuss two different proofs of Vakil's characterisation of smoothable maps to projective space: one is based on Hu-Li's local equations; for the other one we give a classification of genus $1$ singularities.

In \S~\ref{section:p-fields} we adapt Chang-Li's work on $p$-fields to our situation: we introduce the moduli space $\Mone{1}{\PP^4}{d}^p$ of $1$-stable maps with $p$-fields, we endow it with a cosection-localised $0$-dimensional virtual class supported on a proper substack (depending on a homogeneous polynomial $\w$), and we show that its degree coincides with the genus $1$ cuspidal invariants of the quintic threefold $X_5=V(\w)$ up to a sign.

In \S~\ref{sec:comparison} we argue in two different ways that there is a morphism:
 \[\mathfrak M_{1,n}^{\operatorname{wt},\text{st}}\to\mathfrak M_{1,n}^{\operatorname{wt},\text{st}}(1)\]
extending the identity on smooth elliptic curves and replacing elliptic tails of weight $0$ by cusps. We then use it to define:
 \[\Z:= \mathfrak M_{1,n}^{\operatorname{wt},\text{st}}\times_{\mathfrak M_{1,n}^{\operatorname{wt},\text{st}}(1)} \oM^{(1)}_{1,n}(\PP^4,d),\quad \Zp:=\mathfrak M_{1,n}^{\operatorname{wt},\text{st}}\times_{\mathfrak M_{1,n}^{\operatorname{wt},\text{st}}(1)} \oM^{(1)}_{1,n}(\PP^4,d)^p.\]
 We show that $\Z$ is a closed substack of $\oM_1(\PP^4,d)$; unfortunately we are not able to compare $\Zp$ directly with $\oM_1(\PP^4,d)^p$.

For this reason in \S\S~\ref{sec:equations} and \ref{section:main_proof} we study local equations and a desingularisation of $\Zp$, which allows us to adjust Chang-Li's splitting of the intrinsic normal cone to our case. Such analysis allows us to prove the main theorem.

\subsection{Notations and conventions} We work over an algebraically closed field $\k$ of characteristic $0$.

Throughout the paper we shall have fixed a homogeneous polynomial $\mathbf w\in\k[x_0,\ldots,x_4]_5$ of degree $5$ such that $X_5=V(\mathbf w)\subseteq\PP^4$ is a generic smooth quintic threefold. We usually drop the subscript and write $X$ for $X_5$. Furthermore let us consider a fixed positive integer $d$ determining all the homology classes $\beta=d[\ell]\in A_1(\PP^4)$, line bundle weights, etc. that we shall take into account.

@@

Let $\MM_{1,n}$ denote the algebraic stack of prestable $n$-marked curves of arithmetic genus $1$. In the following definitions, by \emph{special point} we mean the preimage of a node or a marking in the normalisation.
\begin{dfn}
Let $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$ be the stack of \emph{weighted-stable} curves of genus $1$ with $n$ markings and total weight $d$: geometric points of $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$ represent connected, reduced, nodal, projective curves of arithmetic genus $1$ with $n$ distinct smooth markings and an integer-valued function $\rm{wt}$ on the dual graph. Such a function assumes only nonnegative values, the sum of which on all the vertices of the dual graph is $d$; $\rm{wt}$ is compatible with the specialisation maps, and we further impose the following stability condition: every $p_a=0$ component of weight $0$ has at least three special points (every $p_a=1$ component of weight $0$ has at least one special point).
\end{dfn}
There is an \'{e}tale, non-separated morphism $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}\to\mathfrak M_{1,n}$.
The stability condition is such that the forgetful map $\M{1}{n}{\PP^r}{d}\to\mathfrak M_{1,n}$ factors through $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$, the weight coming from the degree of the map to $\PP^r$.
We shall denote by $\pi\colon \cC\to \MM_{1,n}$ the universal curve (and use similar notation for its pullbacks).
We shall also consider the universal Picard stack $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}$ of $\pi\colon \cC\to \MM_{1,n}$ with universal line bundle $\mathcal{L}$, subject to the stability condition that $\omega_\pi^{\rm{log}}\otimes\mathcal L^{\otimes 2}$ is $\pi$-ample, and the requirement that the total degree of $\mathcal L$ on the fibers of $\cC$ is $d$. Let $\PtoM$ denote the projection $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}\to \mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$, where the weight comes from the degree of the line bundle. The forgetful map $\rm M=\M{1}{n}{\PP^r}{d}\to\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$ factors through $\PtoM$ by declaring $\mathcal L_{\rm M}:=f^*\OO_{\PP^r}(1)$. We shall often use shorthand notation $\MM$ for $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}$ and $\mathfrak P$ for $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}$.
\begin{dfn}
Let $\mathfrak M_{1,n}^{\operatorname{wt}=d,\text{st}}(1)$ be the stack of \emph{at worst cuspidal} connected, reduced, $n$-marked, projective curves of arithmetic genus $1$ that are weighted-stable with total weight $d$, i.e. the weight is nonnegative, and every $p_a=0$ component of weight $0$ has at least three special points (every $p_a=1$ component of weight $0$ has at least \emph{two} special points).
\end{dfn}
\begin{remark}
 It follows from the miniversal deformation of the cusp that being at worst cuspidal (i.e. having only nodes and cusps as singolarity) is an open condition on the base of any family of curves.
\end{remark}
We shall denote by $\hat{\pi}\colon \widehat{\cC}\to \MM_{1,n}(1)$ the universal curve (and use similar notation for its pullbacks). We shall also consider the universal Picard stack $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}(1)$ of $\widehat{\cC}\to \mathfrak M_{1,n}(1)$ with universal line bundle $\widehat{\mathcal{L}}$ subject to similar stability conditions to the one spelled out for nodal curves. We shall often use shorthand notation $\widehat{\MM}$ for $\mathfrak M_{1,n}^{\rm{wt}=d,\rm{st}}(1)$ and $\widehat{\mathfrak P}$ for $\mathfrak{Pic}^{\rm{tot deg}=d,\rm{st}}_{1,n}(1)$. Let $\hat{\PtoM}$ denote the projection $\widehat{\mathfrak P}\to \widehat{\MM}$.
